### 2024-04-10 texts/paper_closing.tex
update text sent to olav/alex some days ago

% Created 2024-04-04 do 20:38
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

@@

\setlength{\parindent}{1.2cm}
\author{Johannes }
\date{\today}
\title{}
\hypersetup{
 pdfauthor={Johannes },
 pdftitle={},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 29.2 (Org mode 9.6.21)}, 
 pdflang={English}}
\begin{document}

\section*{Identity and the closure of private art museums}

\subsection*{Introduction}

@@

A number of findings highlight the role of identity in the process of private museum closure. 
First, I find that museums which are recognized by an international museum database are less likely to close than those which are not.
Secondly, the results furthermore indicate that museums with more ambiguous names are more likely to close.
Both these findings possibly indicate that adherence to museum standards enhances legitimacy, which in turn improves survival prospects.
Finally, founder reputation, in particular its loss, is found to be substantial as museums whose founders are dropped from a collector ranking are more likely to close than museums whose founders both are and are not included.

\bigbreak
\noindent
\textbf{keywords}: museum, closure, identity

\subsection*{Theoretical framework}

\subsubsection*{Identity and the survival prospects of non-profit organizations}

An identity that is legitimate (understandable) and favorable (positive) has been shown to have a vast number of positive outcomes for organizations \parencite{Lange_Lee_Dai_2010_reputation}.
As gaining the approval of different stakeholders is necessary for being understood and/or supported by other organizations in the environment, having a clearly defined and favorable identity influences resource acquisition and thereby survival prospects \parencite{Rao_1994_reputation}.
Even for non-profit organizations such as museums, which do not produce tangible goods (a domain where identity-related effects are present, e.g. \cite{Hsu_2015_granted,Bogaert_etal_2014_ecological}), a clear and positive identity might be helpful in a number of ways:
In particular in the context of private museums, decline of legitimacy or reputation could lead to lower visitor numbers, less discounts from art dealers, less government grants or lower chances on collaborations with other museums or corporations, resulting in lower revenues and higher acquisition costs.

The empirical record of the link between identity and survival for non-profit organizations is mixed: 
For example \textcite{Bielefeld_1994_survival} finds that NPOs who pursue less legitimation strategies (obtaining endorsements, lobbying or contributing to local causes) are more likely to close.
However, other studies find limited impact of legitimacy on survival, both when measured via density \parencite{Bogaert_etal_2014_ecological} or from archival sources and interviews \parencite{Fernandez_2007_dissolution}.
Thus, even while this study cannot conclusively answer the overall link between idenity and survival, it nevertheless contributes insights into which dimensions of organizational identity do, and which do not contribute to survival in a particular organizational population; namely that of private art museums.

Legitimacy might be obtained by isomorphism, i.e. adopting features associated with blueprints \parencite{diMaggio_1983_iron}.
Understandability \parencite{Glynn_Abzug_2002_names} might thus be limited if an organization is atypical \parencite{Rosch_1975_family}, i.e. displays an unusual combination of features.
In the case of private museums, such cases might be combinations of features from house museums (e.g. accessibility only by appointment) with full-service museums (e.g. extensive activities). 
Next to "hard" organizational features, adherence to naming conventions has been found to enhance legitimacy \parencite{Glynn_Abzug_2002_names}; organizations that adhere to field norms about name length, name ambiguity (usage of artificial names) and name domain specificity (mentioning industry) were judged more legitimate. 
However, the category of museums could be relatively flexible (for example, it is generally not subject to state regulation), which might result in a relatively high tolerance of diversity as "anything goes" and hence limited "devaluation" of non-conforming, atypical members \parencite{Bogaert_etal_2014_ecological}.

@@

\subsubsection*{Other predictors of organizational survival}

\paragraph*{Age}

@@

\paragraph*{Organizational transformation}

\textcite{Carroll_Khessina_2019_demography} argue that transformation of core features of organizations such as changes in technology or authority can have divergent consequences. 
On the one hand, it can disrupt internal routines as well as external customer relations by decreasing (perceived) reliability and accountability as the change upsets established perceptions, which can lead to higher mortality. 
On the other hand, organizational transformation can be necessary to adapt to a changing environment, and hence be beneficial for survival.

In the case of private museums, the death of the founder might constitute a substantial transformation as authority has to be reconfigured.
Abandonment by insiders (such as divorce of directors) has been argued to contribute to NPO closure \parencite{Duckles_Hager_Galaskiewicz_2005_close}.\footnote{Conflict is similarly argued to contribute to closure but is not testable with the current data.}
Even if plans have been made for a handover, the new leaders might not share the same commitment to art as the original founder, potentially decreasing museum sustainability. 
In the case of private museums, founder death has been speculated to pose a challenge to their sustainability as "seldom do heirs share a similar passion or wish to take on the financial burden of maintaining private museum indefinitely" \parencite[p.234]{Walker_2019_collector}.
However, existing research has not found a straightforward effect of founder death on museum closure \parencite{Velthuis_Gera_forthcoming_fragility,Velthuis_etal_2023_boom}.
Nevertheless as previous research has relied primarily on descriptive statistics, investigating founder death in a multivariate survival model allows to investigate the effect of founder death with higher precision.
\bigbreak
\noindent

\subsection*{Data and Methods}

\subsubsection*{Dependent Variable}

A newly constructed database, assembled from different sources,  is used to measure the survival of private museums (cf. \textcite{Velthuis_etal_2023_boom} for details).
First, existing databases of private art museums are harmonized.
Secondly, the online presences of 14 English art publications were exhaustively searched for any mention of private museums.
After identifying unique institutions, data on these organizations was collected such as opening year, closing year (of those which closed) and founder information.
Currently, the database includes 547 museums (of which 446 are currently open, the remainder having closed or transformed into other organizational forms).
Using this database allows to reconstruct the lifecourse of the each institution with the the unit of analysis as museum-year:
The main dependenent variable is thus an indicator of museum closure, which takes the value of 1 for the museums which close in the year of their closure, and 0 otherwise.
Typically for survival data, museums that do not close during the observation period are right-censored, i.e. for such museums the dependent variable is zero for all years.

\subsubsection*{Main Predictors: Identity Measures}

The organizational identity of a museum is measured primarily in terms of its legitimacy, i.e. its adherence to museum standards, and reputation of its founders.

\paragraph*{Legitimacy}

@@

\paragraph*{Reputation}

Reputation is measured via inclusion of the founder in the Artnews top 200 collector ranking \parencite{Artnews_ranking}.
This ranking, established in 1990, lists each year the 200 collectors the editors of the magazine consider most relevant.
For each year, a museums can have one of three values: "Not included"; in which case the founder is not included in the year in question and has not been so in the past, "Included"; when the founder is included, or "Dropped", for cases where the founder was included at some point in the past but is no longer included in the current year.
Cases where multiple individuals are involved in the founding of a private museum are resolved by aggregating the Artnews ranking inclusion history to the founder founder level (i.e., if one of two founders is dropped from the Artnews ranking but the other remains included, the founder couple as a whole is considered included).

\subsubsection*{Control variables}

\bigbreak
\noindent
A number of control variables are used: 
On the level of the founder, gender constitutes a time-invariant variable; with possible values being male, female and couple.
Organizational transformation, such as change of authority has been to increase chances of closure both generally \parencite{Carroll_Khessina_2019_demography} and in the case of private museums specifically (\cite[p.234]{Walker_2019_collector}: "seldom do heirs share a similar passion or wish to take on the financial burden of maintaining private museum indefinitely").
Therefore, a measure of founder death is constructed as a time-varying covariate, and is set to 1 for museum-years after the death of the founder. 
Centrality of the founder's vision is measured via the proxy of whether the name of the founder is included in the name of the museum.

Furthermore, organizational theory \parencite{hannan89_organ} predicts a U-shaped relationship between density and mortality:
While at low levels, increases in the number of organizations are argued to primarily increase legitimacy and lower exit rates, increases at higher numbers increase the competition over resources and thereby raise mortality rates.
On the level of the environment, therefore the density (number) of private museum per country is included both in linear and squared form.

\subsubsection*{Analytical framework}

First, we investigate the relationship between age and closing descriptively using hazard rates and the Kaplan-Meier survival function.
This non-parametric method estimates the hazard rate as the conditional probability of exiting at time t\textsubscript{i} given being alive as \(h(t_i) = \frac{d_i}{n_i}\), (with \(n_i\) as the number of units at risk at time \(t_i\) and \(d_i\) as the number of exits at time \(t_i\)), and the survival function as the product of these conditional probabilities as \(S(t) = \prod_{t_i \geq t} \left(1-h(t_i) \right)\).

@@

\subsection*{Results}

\subsubsection*{Summary statistics}

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Thu Apr  4 20:37:25 2024
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrrr}
  \hline
 & & \multicolumn{2}{c}{Museum} & \multicolumn{4}{c}{Museum-year} \\ 
\cmidrule(r){3-4}\cmidrule(r){5-8} \multicolumn{1}{l}{} & \multicolumn{1}{l}{Variable} & \multicolumn{1}{l}{Count} & \multicolumn{1}{l}{Mean} & \multicolumn{1}{l}{Mean} & \multicolumn{1}{l}{SD} & \multicolumn{1}{l}{Min.} & \multicolumn{1}{l}{Max.}\\ 
 \hline
  \multicolumn{8}{l}{\textbf{Founder}} \\ 
 & Gender - Male & 275 & 0.574 &  0.593 &  0.49 & 0 & 1 \\ 
   & Gender - Female & 76 & 0.159 &  0.169 &  0.37 & 0 & 1 \\ 
   & Gender - Couple & 128 & 0.267 &  0.238 &  0.43 & 0 & 1 \\ 
   & Founder died & 58 & 0.121 &  0.098 &  0.30 & 0 & 1 \\ 
   \multicolumn{8}{l}{\textbf{Museum}} \\ 
 & Self-Identification - Museum & 178 & 0.372 &  0.410 &  0.49 & 0 & 1 \\ 
   & Self-Identification - Foundation & 101 & 0.211 &  0.189 &  0.39 & 0 & 1 \\ 
   & Self-Identification - Collection & 74 & 0.154 &  0.140 &  0.35 & 0 & 1 \\ 
   & Self-Identification - Other & 126 & 0.263 &  0.261 &  0.44 & 0 & 1 \\ 
   & Founder name in Museum name & 229 & 0.478 &  0.488 &  0.50 & 0 & 1 \\ 
   & MOW inclusion & 91 & 0.190 &  0.304 &  0.46 & 0 & 1 \\ 
   & AN Ranking - Not Included &  &  &  0.797 &  0.40 & 0 & 1 \\ 
   & AN Ranking - Included &  &  &  0.150 &  0.36 & 0 & 1 \\ 
   & AN Ranking - Dropped &  &  &  0.054 &  0.23 & 0 & 1 \\ 
   \multicolumn{8}{l}{\textbf{Environment}} \\ 
 & PM density &  &  & 20.889 & 18.49 & 1 & 62 \\ 
   & Pop. (millions) within 10km &  &  &  1.378 &  1.74 & 0.000087 & 10.81 \\ 
   & Nbr PM within 10km &  &  &  2.707 &  2.99 & 1 & 16 \\ 
   & Europe and North America & 312 & 0.651 &  0.645 &  0.48 & 0 & 1 \\ 
   & Region - Africa & 7 & 0.015 &  0.012 &  0.11 & 0 & 1 \\ 
   & Region - Asia & 128 & 0.267 &  0.282 &  0.45 & 0 & 1 \\ 
   & Region - Europe & 241 & 0.503 &  0.501 &  0.50 & 0 & 1 \\ 
   & Region - Latin America & 22 & 0.046 &  0.042 &  0.20 & 0 & 1 \\ 
   & Region - North America & 71 & 0.148 &  0.143 &  0.35 & 0 & 1 \\ 
   & Region - Oceania & 10 & 0.021 &  0.019 &  0.14 & 0 & 1 \\ 
   \hline
\end{tabular}
\caption{Summary Statistics} 
\label{tbl:t_sumstats}
\end{table}

@@

\subsubsection*{Hazard rate and Age Dependence}

\begin{figure}[htbp]
\centering
\includegraphics[width=16cm]{../figures/p_hazard.pdf}
\caption{\label{fig:p_hazard}Private Museum hazard function}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=14cm]{../figures/p_surv.pdf}
\caption{\label{fig:p_surv}Private Museum Survival probability}
\end{figure}

@@

Overall the entire life span, the average hazard rate is 0.78\%, which allows comparison to other studies of NPO closure.
While after very cursory search of the literature I did not find a systematic literature review or meta-analysis, a number of individual studies had investigated similar populations (see table \ref{tbl:litreview}).

\begin{table}[htbp]
\caption{\label{tbl:litreview}Hazard rates in studies of museums and other non-profits}
\centering
\begin{tabular}{llll}
\hline
study & population & avg. hazard rate & data source\\
\hline
\cite{Hager_2001_vulnerability} & Art Museums & 2.4\% & IRS\\
\cite{Gordon_etal_2013_insolvency} & Museums & 0.7\% & IRS\\
\hline
\cite{Hager_2001_vulnerability} & all Art NPOs & 2.8\% & IRS\\
\cite{Gordon_etal_2013_insolvency} & all Art NPOs & 2.08\% & IRS\\
\cite{Gordon_etal_2013_insolvency} & all NPOs & 1.58\% & IRS\\
\cite{Hager_Galaskiewicz_Larson_2007_liability} & NPOs & 1.25\% & own data\\
\cite{Clifford_2018_reinforcing} & Charities & 2\%-5\% & Reg. of Charities (UK)\\
\cite{Mayer_2022_slimmer} & NPOs & 0.17\% & IRS\\
\hline
\end{tabular}
\end{table}

\textcite{Hager_2001_vulnerability} and \textcite{Gordon_etal_2013_insolvency} both study NPOs more generally, but report survival estimates by organizational form.
At the moment it is not clear which factors account for their hazard rates to differ substantially, but it might be due to selection criteria.
Both these studies also report substantial variation in hazard rates between organizational forms.
While preliminary results indicate that private museums have dissolution rates lower or similar to museums and non-profit organizations generally, more research is needed to establish the comparability of these hazard rates.

\subsubsection*{Regression Results}

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Thu Apr  4 20:37:33 2024
\begin{table}[ht]
\centering
\begin{tabular}{p{0mm}lD{)}{)}{8)3}}
  \hline 
 \multicolumn{1}{l}{} & \multicolumn{1}{l}{Variable} & \multicolumn{1}{l}{r\_more}\\ 
 \hline
  \multicolumn{3}{l}{\textbf{Founder}} \\ 
 & Gender - Female & .20 \; (0.36) \\ 
   & Gender - Couple & .12 \; (0.37) \\ 
   & Founder died & .21 \; (0.54) \\ 
   \multicolumn{3}{l}{\textbf{Museum}} \\ 
 & Self-Identification - Foundation & .84 \; (0.48) \\ 
   & Self-Identification - Collection & .46 \; (0.58) \\ 
   & Self-Identification - Other & 1.34 \; (0.38)^{***} \\ 
   & AN Ranking - Included & .64 \; (0.38) \\ 
   & AN Ranking - Dropped & .87 \; (0.53) \\ 
   & Founder name in Museum name & -.27 \; (0.33) \\ 
   & MOW inclusion & -.84 \; (0.40)^{*} \\ 
   \multicolumn{3}{l}{\textbf{Environment}} \\ 
 & PM density & -.05 \; (0.03) \\ 
   & PM density$^{2}$ & .00 \; (0.00) \\ 
   & Pop. (millions) within 10km & .02 \; (0.11) \\ 
   & Nbr PM within 10km & .12 \; (0.06)^{*} \\ 
   \hline
 & Museum-years & 6423 \\ 
   & Closures & 52 \\ 
   & log. Likelihood & -257.05 \\ 
   & AIC & 542.09 \\ 
   & BIC & 569.41 \\ 
   \hline 
 \multicolumn{3}{l}{\footnotesize{standard errors in parantheses.\textsuperscript{***}p $<$ 0.001;\textsuperscript{**}p $<$ 0.01;\textsuperscript{*}p $<$ 0.05.}}
\end{tabular}
\caption{Cox Proportional Hazards Regression Results} 
\label{tbl:t_reg_coxph}
\end{table}

>>>>

% Created 2024-04-07 zo 17:21
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

@@

\setlength{\parindent}{1.2cm}
\author{Johannes Aengenheyster}
\date{\today}
\title{}
\hypersetup{
 pdfauthor={Johannes Aengenheyster},
 pdftitle={},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 29.2 (Org mode 9.6.21)}, 
 pdflang={English}}
\begin{document}

\section*{Identity and the closure of private art museums}
\label{sec:org16a7685}

\subsection*{Introduction}
\label{sec:orgef6a080}

@@

\subsection*{Theoretical framework}
\label{sec:org90170ed}

As a substantial literature on organizational survival exists,

\subsubsection*{Identity and the survival prospects of non-profit organizations}
\label{sec:org8725d02}
An identity that is legitimate (understandable) and favorable (positive) has been shown to have a vast number of positive outcomes for organizations \parencite{Lange_Lee_Dai_2010_reputation}.
As gaining the approval of different stakeholders is necessary for being understood and/or supported by other organizations in the environment, having a clearly defined and favorable identity influences resource acquisition and thereby survival prospects \parencite{Rao_1994_reputation}.
Even for non-profit organizations such as museums, which do not produce tangible goods (a domain where identity-related effects are present, e.g. \cite{Hsu_2015_granted,Bogaert_etal_2014_ecological}), a clear and positive identity might be helpful in a number of ways:
In particular in the context of private museums, decline of legitimacy or reputation could lead to lower visitor numbers, less discounts from art dealers, less government grants or lower chances on collaborations with other museums or corporations, resulting in lower revenues and higher acquisition costs.

The empirical record of the link between identity and survival for non-profit organizations is mixed: 
For example \textcite{Bielefeld_1994_survival} finds that NPOs who pursue less legitimation strategies (obtaining endorsements, lobbying or contributing to local causes) are more likely to close.
However, other studies find limited impact of legitimacy on survival, both when measured via density \parencite{Bogaert_etal_2014_ecological} or from archival sources and interviews \parencite{Fernandez_2007_dissolution}.
Legitimacy might be obtained by isomorphism, i.e. adopting features associated with blueprints \parencite{diMaggio_1983_iron}.
Understandability \parencite{Glynn_Abzug_2002_names} might thus be limited if an organization is atypical \parencite{Rosch_1975_family}, i.e. displays an unusual combination of features.
In the case of private museums, such cases might be combinations of features from house museums (e.g. accessibility only by appointment) with full-service museums (e.g. extensive activities). 
Next to "hard" organizational features, adherence to naming conventions has been found to enhance legitimacy \parencite{Glynn_Abzug_2002_names}; organizations that adhere to field norms about name length, name ambiguity (usage of artificial names) and name domain specificity (mentioning industry) were judged more legitimate. 
However, the category of museums could be relatively flexible (for example, it is generally not subject to state regulation), which might result in a relatively high tolerance of diversity as "anything goes" and hence limited "devaluation" of non-conforming, atypical members \parencite{Bogaert_etal_2014_ecological}.

@@

\subsubsection*{Organizational transformation}
\label{sec:org19d4baf}
\textcite{Carroll_Khessina_2019_demography} argue that transformation of core features of organizations such as changes in technology or authority can have divergent consequences. 
On the one hand, it can disrupt internal routines as well as external customer relations by decreasing (perceived) reliability and accountability as the change upsets established perceptions, which can lead to higher mortality. 
On the other hand, organizational transformation can be necessary to adapt to a changing environment, and hence be beneficial for survival.

In the case of private museums, the death of the founder might constitute a substantial transformation as authority has to be reconfigured.
Abandonment by insiders (such as divorce of directors) has been argued to contribute to NPO closure \parencite{Duckles_Hager_Galaskiewicz_2005_close}.\footnote{Conflict is similarly argued to contribute to closure but is not testable with the current data.}
Even if plans have been made for a handover, the new leaders might not share the same commitment to art as the original founder, potentially decreasing museum sustainability. 
In the case of private museums, founder death has been speculated to pose a challenge to their sustainability as "seldom do heirs share a similar passion or wish to take on the financial burden of maintaining private museum indefinitely" \parencite[p.234]{Walker_2019_collector}.
However, existing research has not found a straightforward effect of founder death on museum closure \parencite{Velthuis_Gera_forthcoming_fragility,Velthuis_etal_2023_boom}.
Nevertheless as previous research has relied primarily on descriptive statistics, investigating founder death in a multivariate survival model allows to investigate the effect of founder death with higher precision.
\bigbreak
\noindent
\textbf{Hypothesis 2}: A private museum is more likely to close after the death of the founder.

\subsubsection*{Competition}
\label{sec:org9840956}

needs to write out: 
\begin{itemize}
\item \cite{Hager_1999_demise}:
\begin{itemize}
\item p.173: Ten organizations cited an inability to compete as a proximate cause for closure, making it the most popular proximate cause of organizational closure.
\item p.173: Both [art organizations] noted that the majority of the community’s arts dollars, both individual donations/subscriptions and community and corporate grants, tended to go to the singular, most reputable arts organizations of their types (e.g. museums, orchestras, dance troupes) in the metro area. Other arts organizations faced strong competitive pressures from these flagship organizations, a factor which caused the demise of Organizations 276 and 293
\end{itemize}
\item \cite{Lecy_2010_nonprofit}:
\begin{itemize}
\item p.27: The analysis shows that a period of rapid sector expansion – a period where many additional resources are being poured into the sector – creates a more competitive environment as evidenced by higher rates of organizational mortality.
\end{itemize}
\item \cite{Park_Shon_Lu_2021_mortality}
\begin{itemize}
\item p.844: Furthermore, nonprofits that run their business in a county with greater organizational density and population size are more likely to dissolve, which implies that competition, as well as demand for service, prevents nonprofits from surviving
\end{itemize}
\item \cite{HernandezOrtiz_2022_discontinuity}
\begin{itemize}
\item p.116:  A crowded niche increased the competition and limited their access to resources and clients. As reported by informant of organization \#C07: "Well, there was less money to go around and more organizations seeking the money. Here we have the same, the same group of people that were supporting the organization, you know, donating to it and that only goes so far. So, we had a full-time grant writer, we were always applying for grants. But there are so many organizations applying for the same grants". (P31, Pos. 49-51)
\end{itemize}
\item \cite{Lu_Shon_Zhang_2019_dissolution}
\begin{itemize}
\item p.43: For example, nonprofits operating in a more competitive environment with a higher degree of organizational density would experience a higher chance of dissolution.
\end{itemize}
\item \cite{Haugh_etal_2021_nascent}:
\begin{itemize}
\item p.1227: SEs are likely to compete with non-profit organizations for grants,donations, and programme-related investments. Second,both SEs and non-profit organizations serve the disadvantaged and socially excluded and thus may also compete for service users. Although collaboration between SEs and charities could increase the supply of services to those in need (Smiddy 2010), they compete with each other for resources and beneficiaries. The more similar the resource needs and markets served, the greater the competition between the organizational forms (Baum and Haveman 1997). Hence, the population density ofregistered non-profit organizations will thus have a detrimental effect on SE survival.
\end{itemize}
\end{itemize}

\textbf{Hypothesis 3}: A private museum is more likely to close in a more competitive environment.

\subsubsection*{Other predictors of organizational survival}
\label{sec:org9672c62}
\paragraph*{Age}
\label{sec:orgae2e32a}

@@

\subsection*{Data and Methods}
\label{sec:org6923afe}

\subsubsection*{Dependent Variable}
\label{sec:org7efae3c}
A newly constructed database, assembled from different sources,  is used to measure the survival of private museums (cf. \textcite{Velthuis_etal_2023_boom} for details).
First, existing databases of private art museums are harmonized.
Secondly, the online presences of 14 English art publications were exhaustively searched for any mention of private museums.
After identifying unique institutions, data on these organizations was collected such as opening year, closing year (of those which closed) and founder information.
Currently, the database includes 547 museums (of which 446 are currently open, the remainder having closed or transformed into other organizational forms).
Using this database allows to reconstruct the lifecourse of the each institution with the the unit of analysis as museum-year:
The main dependenent variable is thus an indicator of museum closure, which takes the value of 1 for the museums which close in the year of their closure, and 0 otherwise.
Typically for survival data, museums that do not close during the observation period are right-censored, i.e. for such museums the dependent variable is zero for all years.

\subsubsection*{Main Predictors}
\label{sec:orge0e80eb}

\paragraph*{Legitimacy}
\label{sec:org9952eec}

@@

\paragraph*{Reputation}
\label{sec:org051c5db}

Reputation is measured via inclusion of the founder in the Artnews top 200 collector ranking \parencite{Artnews_ranking}.
This ranking, established in 1990, lists each year the 200 collectors the editors of the magazine consider most relevant.
For each year, a museums can have one of three values: "Not included"; in which case the founder is not included in the year in question and has not been so in the past, "Included"; when the founder is included, or "Dropped", for cases where the founder was included at some point in the past but is no longer included in the current year.
Cases where multiple individuals are involved in the founding of a private museum are resolved by aggregating the Artnews ranking inclusion history to the founder level (i.e., if only one of two founders is dropped from the Artnews ranking but the other remains included, the founder couple as a whole is considered included).
The variable is lagged by one year to avoid reverse causality, i.e. to exclude the possibility that an association is observed which is present due to founders being dropped from the ranking as a consequence of the closure of their museum.

\paragraph*{Organizational transformation}
\label{sec:org671fa45}
A measure of founder death is constructed as a time-varying covariate, and is set to 1 for museum-years after the death of the founder.

\paragraph*{Competition}
\label{sec:orgc059910}
On a national level, density of private museum is calculated as the number of private museums per million.
On a local level, I use the Global Human Settlement Layer \parencite{EC_2023_GHSL} to estimate the number of private museums and population counts inside a radius of ten kilometers around each private museum for each year.
To be able to estimate both competition and demand/audience/population separately, I do not calculate densities as per capita rates, but include the local museum counts, local population in millions as well as an interaction between the two.

\subsubsection*{Control variables}
\label{sec:org4412539}

\bigbreak
\noindent
A number of control variables are used: 
On the level of the founder, gender constitutes a time-invariant variable; with possible values being male, female and couple.
Centrality of the founder's vision is measured via the proxy of whether the name of the founder is included in the name of the museum.

\subsubsection*{Analytical framework}
\label{sec:orge4889f4}

First, I investigate the relationship between age and closing descriptively using hazard rates and the Kaplan-Meier survival function.
This non-parametric method estimates the hazard rate as the conditional probability of exiting at time t\textsubscript{i} given being alive as \(h(t_i) = \frac{d_i}{n_i}\), (with \(n_i\) as the number of units at risk at time \(t_i\) and \(d_i\) as the number of exits at time \(t_i\)), and the survival function as the product of these conditional probabilities as \(S(t) = \prod_{t_i \geq t} \left(1-h(t_i) \right)\).

@@

\subsection*{Results}
\label{sec:org9e38ff3}

\subsubsection*{Summary statistics}
\label{sec:org12efa6c}

% latex table generated in R 4.3.3 by xtable 1.8-4 package
% Sun Apr  7 12:05:30 2024
\begin{table}[ht]
\centering
\begin{tabular}{llrrrrrr}
  \hline
 & & \multicolumn{2}{c}{Museum} & \multicolumn{4}{c}{Museum-year} \\ 
\cmidrule(r){3-4}\cmidrule(r){5-8} \multicolumn{1}{l}{} & \multicolumn{1}{l}{Variable} & \multicolumn{1}{l}{Count} & \multicolumn{1}{l}{Mean} & \multicolumn{1}{l}{Mean} & \multicolumn{1}{l}{SD} & \multicolumn{1}{l}{Min.} & \multicolumn{1}{l}{Max.}\\ 
 \hline
  \multicolumn{8}{l}{\textbf{Founder}} \\ 
 & Gender - Male & $275$ & $0.574$ & $  0.593$ & $  0.49$ & $0$ & $1$ \\ 
   & Gender - Female & $76$ & $0.159$ & $  0.169$ & $  0.37$ & $0$ & $1$ \\ 
   & Gender - Couple & $128$ & $0.267$ & $  0.238$ & $  0.43$ & $0$ & $1$ \\ 
   & Founder died & $58$ & $0.121$ & $  0.098$ & $  0.30$ & $0$ & $1$ \\ 
   \multicolumn{8}{l}{\textbf{Museum}} \\ 
 & Self-Identification - Museum & $178$ & $0.372$ & $  0.410$ & $  0.49$ & $0$ & $1$ \\ 
   & Self-Identification - Foundation & $101$ & $0.211$ & $  0.189$ & $  0.39$ & $0$ & $1$ \\ 
   & Self-Identification - Collection & $74$ & $0.154$ & $  0.140$ & $  0.35$ & $0$ & $1$ \\ 
   & Self-Identification - Other & $126$ & $0.263$ & $  0.261$ & $  0.44$ & $0$ & $1$ \\ 
   & Founder name in Museum name & $229$ & $0.478$ & $  0.488$ & $  0.50$ & $0$ & $1$ \\ 
   & MOW inclusion & $91$ & $0.190$ & $  0.304$ & $  0.46$ & $0$ & $1$ \\ 
   & AN Ranking - Not Included & $NA$ & $NA$ & $  0.797$ & $  0.40$ & $0$ & $1$ \\ 
   & AN Ranking - Included & $NA$ & $NA$ & $  0.150$ & $  0.36$ & $0$ & $1$ \\ 
   & AN Ranking - Dropped & $NA$ & $NA$ & $  0.054$ & $  0.23$ & $0$ & $1$ \\ 
   \multicolumn{8}{l}{\textbf{Environment}} \\ 
 & PM density (country) & $NA$ & $NA$ & $  0.411$ & $  1.23$ & $0.00077$ & $27.26$ \\ 
   & Pop. (millions) within 10km & $NA$ & $NA$ & $  1.322$ & $  1.69$ & $0.000087$ & $10.81$ \\ 
   & Pop. (millions) country & $NA$ & $NA$ & $156.197$ & $287.82$ & $0.037$ & $1412.36$ \\ 
   & Nbr PM within 10km & $NA$ & $NA$ & $  1.601$ & $  2.97$ & $0$ & $15$ \\ 
   & PM density (10km) & $NA$ & $NA$ & $  1.423$ & $  4.01$ & $0$ & $41.85$ \\ 
   & Europe and North America & $312$ & $0.651$ & $  0.645$ & $  0.48$ & $0$ & $1$ \\ 
   & Region - Africa & $7$ & $0.015$ & $  0.012$ & $  0.11$ & $0$ & $1$ \\ 
   & Region - Asia & $128$ & $0.267$ & $  0.282$ & $  0.45$ & $0$ & $1$ \\ 
   & Region - Europe & $241$ & $0.503$ & $  0.501$ & $  0.50$ & $0$ & $1$ \\ 
   & Region - Latin America & $22$ & $0.046$ & $  0.042$ & $  0.20$ & $0$ & $1$ \\ 
   & Region - North America & $71$ & $0.148$ & $  0.143$ & $  0.35$ & $0$ & $1$ \\ 
   & Region - Oceania & $10$ & $0.021$ & $  0.019$ & $  0.14$ & $0$ & $1$ \\ 
   \hline
\end{tabular}
\caption{Summary Statistics} 
\label{tbl:t_sumstats}
\end{table}

@@

\subsubsection*{Hazard rate and Age Dependence}
\label{sec:org01e6edd}

\begin{figure}[htbp]
\centering
\includegraphics[width=16cm]{../figures/p_hazard.png}
\caption{\label{fig:p_hazard}Private Museum hazard function}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=14cm]{../figures/p_surv.png}
\caption{\label{fig:p_surv}Private Museum Survival probability}
\end{figure}

@@

Overall the entire life span, the average hazard rate is 0.78\%, which allows comparison to other studies of NPO closure.
While I did not find a systematic literature review or meta-analysis, a number of individual studies had investigated similar populations based on IRS data (see table \ref{tbl:litreview})

\begin{table}[htbp]
\caption{\label{tbl:litreview}Hazard rates in studies of museums art Art NPOs}
\centering
\begin{tabular}{llll}
\hline
study & population & avg. hazard rate & org years\\
\hline
\cite{Hager_2001_vulnerability} & Art Museums & 2.4\% & \\
\cite{Gordon_etal_2013_insolvency} & Museums & 0.7\% & \\
\hline
\cite{Hager_2001_vulnerability} & all Art NPOs & 2.8\% & \\
\cite{Gordon_etal_2013_insolvency} & all Art NPOs & 2.08\% & \\
\hline
\end{tabular}
\end{table}

\textcite{Hager_2001_vulnerability} and \textcite{Gordon_etal_2013_insolvency} both study NPOs more generally, but report survival estimates by organizational form.
Both these studies also report substantial variation in hazard rates between organizational forms.
While these comparisons indicate that private museums have dissolution rates lower or similar to museums and non-profit organizations generally, more research is needed to establish the comparability of these hazard rates.

Comparison to other forms of nonprofit organizations reported by \parencite{Gordon_etal_2013_insolvency}
places private museum at a risk similar to that of animal shelters and zoos (0.7\%), botanical \& environmental centers (0.8\%), colleges and universities (0.8\%), food and agricultural-oriented human service organizations (0.8\%) and  student and educational services (0.8\%).
While the diversity of non-profit organizations makes comparisons to its entirety difficult, it is still worthwhile to consider that the average hazard of private museums 0.78\% is around half that of that NPOs generally (1.58\%, see table \ref{tbl:litreview2} for studies on wider nonprofit populations).
It is furthermore substantially lower than that of youth centers (1.4\%), children-focused human services (2.4\%), K12 schools (1.8\%), mental health organizations (2.4\%) or performing arts organizations (3.7\%).
In fact, lower dissolution rates are only reported for historical societies (0.5\%) and libraries (0.2\%).

\begin{table}[htbp]
\caption{\label{tbl:litreview2}Hazard rates in studies of NPOs}
\centering
\begin{tabular}{llll}
\hline
study & population & avg. hazard rate & data source\\
\hline
\cite{Gordon_etal_2013_insolvency} & NPOs & 1.58\% & IRS\\
\cite{Hager_Galaskiewicz_Larson_2007_liability} & NPOs & 1.25\% & own data\\
\cite{Clifford_2018_reinforcing} & Charities & 2\%-5\% & Reg. of Charities (UK)\\
\cite{Mayer_2022_slimmer} & public charities & 0.17\% & IRS\\
\hline
\end{tabular}
\end{table}

\subsubsection*{Regression Results}
\label{sec:org4878b83}
